\vspace{-2cm} Kronos \& Krios: Evidence for accretion of rocky planets in a comoving pair of G dwarfs \vspace{-0.5cm}

### 2017-06-22 paper/ms.tex
Move things around and add new figures

% Titles:
% - 90 characters (including spaces)
% - do not normally include numbers, acronyms, abbreviations or punctuation
% - They should include sufficient detail for indexing purposes but be general
%   enough for readers outside the field to appreciate what the paper is about.
\title{
\vspace{-2cm}
  Kronos \& Krios: Evidence for accretion of rocky planets in a comoving pair of G dwarfs
  % \vspace{-1cm}
}

\author{
  Semyeong Oh$^{1,*}$,\,
  Adrian M. Price-Whelan$^{1}$,\,
  John M. Brewer$^{2}$,\,\\
  David W. Hogg$^{3,4,5,6}$,\,
  David N. Spergel$^{1,3}$,\,
  Justin Myles$^{2}$
}

@@

% affiliations
{\small\noindent
  $^*$To whom correspondence should be addressed: \texttt{semyeong@astro.princeton.edu} \\
  $^1$Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA \\
  $^2$Department of Astronomy, Yale University, 260 Whitney Ave, New Haven, CT 06511, USA \\
  $^3$Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA \\
  $^4$Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA \\
  $^5$Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA \\
  $^6$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg \\
}

@@

% coevality based on stellar parameters
% TODO reference supplement discussion on age here?
Their stellar parameters also indicate that they are coeval with an age of $\sim 4$~Gyr.
We use the distances (inferred from \gaia\ parallaxes), $V$-band magnitudes,
and $B-V$ colors to obtain bolometric luminosities of the two stars\cite{2003AJ....126..778V}.
We then combine the luminosities with $T_\mathrm{eff}$, \elemH{Fe}, \elemH{Si}
in order interpolate the mass, age, radius, and $\log(g)$ of each star
in a grid of Yale-Yonsei model isochrones\cite{2013ApJ...776...87S}.
The best-fit isochrone ages of \sunanalog\ and \bizarreone\ are
$4.00_{-1.56}^{+1.51}$~Gyr and $4.28_{-1.03}^{+1.11}$~Gyr, respectively,
consistent with them being coeval.

@@

% high Li
Both stars have high surface \elem{Li} abundances considering
their ages of $\sim 4$~Gyr.
The surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning. %TODO: REFERENCE
The \elem{Li} doublet at $6707.6$~\AA, which is usually too weak to be
detectable at this age even for high signal-to-noise spectra, is clearly
visible in the spectra of both stars (see \figname~\ref{fig:spec2}).
The measured absolute abundances for \bizarreone\ and \sunanalog\ are
$2.75$~dex and $2.25$~dex, respectively (Myles \etal, in prep),
resulting in the largest abundance difference ($\approx 0.5$~dex) seen
among all measured elements.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{toycalc.png}
  \caption{
    Comparing the observed abundance difference vs. \Tcondens\ rank
    to the expected change in solar surface abundance after adding $20$~\mearth\ of
    material with bulk Earth composition (\cite{mcdonough2001composition}).
    The assumed mass fraction in the convective zone is $0.03$.
    All metals are ranked by their \Tcondens\ for solar composition gas,
    and the condensation temperature may be read from the gray line and right y-axis,
    same as in \figname~\ref{fig:relabun_tcrank}.
    Locations of the elements measured for \bizarreone-\sunanalog\ pair are
    indicated by a vertical line and its symbol.
    The close match with the observed abundance difference in \bizarreone-\sunanalog\ pair
    suggests that the abundance difference may be due to accretion of
    $20$~\mearth\ of rocky planetary material.
    The element \elem{Li} is off the plot and indicated with a red arrow
    (see text for details).
  }
  \label{fig:toycalc}
\end{figure}

We show the abundance difference between \bizarreone\ and \sunanalog\ ordered
by each element's equilibrium \Tcondens\ in solar system composition
gas\citet{2003ApJ...591.1220L} in Figure~\ref{fig:toycalc}.
The five under-enhanced elements in \bizarreone\
relative to \sunanalog\ are the five most volatile in all elements measured.
The difference in \elem{Mn} ($\Tcondens = 1158$~K) and
\elem{Cr} ($\Tcondens = 1296$~K) suggests a break in $\Tcondens \approx 1200$~K.

We interpret the \Tcondens-dependent metal enhancement of \bizarreone\ relative
to \sunanalog\ and the very high surface \elem{Li} abundance ($A(\elem{Li}) =
2.75$) of \bizarreone\ as evidence for accretion of rocky (refractory-rich)
planetary material in \bizarreone.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering\cite{} or encounters\cite{Malmberg:2011aa} with a field star can
lead some planets to be ejected or accreted onto its host star.
If one of the two stars in a wide binary accreted refractory-rich (rocky) planets
more than the other, the accretion may imprint a strong \Tcondens\-dependent
abundance differences like those observed in \bizarreone-\sunanalog.

@@

We estimate that accretion of $\approx 20$~\mearth\ of rocky planets
in \bizarreone\ relative to \sunanalog\ can explain the enhancement in
refractory elements as well as the surface Li abundance.

where $f_\mathrm{CZ}$ is the fraction of the star's mass in the convective envelope.
We take the composition of bulk Earth from a chondritic model of the Earth
(\cite{mcdonough2001composition}).
Similar calculations have been performed by \citet{Chambers:2010aa} and
\citet{Mack:2014aa,Mack:2016aa}.

Figure~\ref{fig:toycalc} shows the expected change of surface abundances of
metals in a Sun-like star after $20~\mearth$ of material with composition of
bulk Earth is added.
A volatility trend that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known (\cite{mcdonough2001composition}).
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding $20~\mearth$ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggesting that the refractory-enhanced star, \bizarreone\,
accreted $20~\mearth$ more of rocky planetary material than \sunanalog.

@@

%TODO: augment this paragraph
% - more geochemical study references
We stress that while the calculation carried out is useful in
an order-of-magnitude sense, further investigation of each of the simplifying
assumptions made is warranted.
In addition, the composition of bulk Earth has some uncertainties.
For example, the reported bulk Earth concentration of the siderphile element
\elem{Mn}, varies from $800$ to $1700$~ppm (\cite{1998psc..book.....L}).
While the latter value from \cite{mcdonough2001composition} has been used in
our calculation, the former value would bring the observed $\Delta\elemH{Mn}$
to an even closer agreement.
Given these limitations, the level of agreement for $\Delta\elemH{X}$ {\it and}
\elem{Li} for \bizarreone\ is remarkable.

@@

\begin{table*}[htpb]
  \caption{Astrometric and spectroscopic measurements.}
  \label{tab:t2}
  \centering
  \begin{tabular}{ccc}
\hline\hline
Name & HD 240429 & HD 240430 \\
\hline
Sp Type                                   & G2                & G0                \\
$T_\mathrm{eff}$                          & 5878              & 5803              \\
$\log{g}$                                 & 4.43              & 4.33              \\
$v\sin{i}$                                & 1.1               & 2.5               \\
$[\elem{Fe}/\elem{H}]$                    & 0.01              & 0.20              \\
$v_r$                                     & $-21.2$           & $-21.2$           \\
$\varpi$ \footnotemark[1]                 & $9.35 \pm 0.24$   & $9.41 \pm 0.25$   \\
$\mu_\alpha^*$ \footnotemark[1]           & $89.25 \pm 0.66$  & $89.41 \pm 0.69$  \\
$\mu_\delta$ \footnotemark[1]             & $-29.68 \pm 0.54$ & $-30.12 \pm 0.52$ \\
\hline
\multicolumn{3}{c}{$T_c < 1200$~K} \\
\hline
$A(\elem{Li})$ \footnotemark[2]           & $2.25$            & $2.75$            \\
$\elemH{C}$                               & $0.00$            & $0.09$            \\
$\elemH{N}$                               & $-0.06$           & $-0.01$           \\
$\elemH{O}$                               & $0.01$            & $0.09$            \\
$\elemH{Na}$                              & $-0.06$           & $-0.04$           \\
$\elemH{Mn}$                              & $-0.03$           & $0.00$            \\
\hline
\multicolumn{3}{c}{$T_c > 1200$~K} \\
\hline
$\elemH{Mg}$                              & $0.01$            & $0.19$            \\
$\elemH{Al}$                              & $0.01$            & $0.21$            \\
$\elemH{Si}$                              & $0.00$            & $0.16$            \\
$\elemH{Ca}$                              & $0.02$            & $0.23$            \\
$\elemH{Ti}$                              & $0.02$            & $0.20$            \\
$\elemH{V}$                               & $0.02$            & $0.20$            \\
$\elemH{Cr}$                              & $0.01$            & $0.17$            \\
$\elemH{Fe}$                              & $0.01$            & $0.20$            \\
$\elemH{Ni}$                              & $-0.01$           & $0.21$            \\
$\elemH{Y}$                               & $0.04$            & $0.26$            \\
\hline\hline
\end{tabular}
%
%\begin{tablenotes}
%  \item All spectroscopic measurements except $\elem{Li}$ are from \citealt{2016ApJS..225...32B}.
%  \footnotetext[1]{From \tgas. }
%  \footnotetext[2]{Absolute abundance from \todo{CITE}.}
%\end{tablenotes}
\end{table*}

@@

Combining the precise radial velocities obtained from spectra with the \tgas\
astrometry, we can compare differences between the inferred 6D phase-space
coordinates of the two stars.
We start by generating posterior samples over the Heliocentric distance, $r$,
tangential velocities, $(v_\alpha, v_\delta)$, and radial velocity, $v_r$,
given the observed parallax, $\hat\pi$, and proper motions,
$(\hat\mu_{\alpha^*}, \hat\mu_\delta)$, and radial velocity, $\hat v_r$.
We assume the noise is Gaussian, and the radial velocity measurement is
uncorrelated with the astrometric measurements.
If we define
\begin{eqnarray}
  \vec{\hat y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          \hat\pi &
          \hat\mu_{\alpha^*} &
          \hat\mu_\delta &
          \hat v_r
        \end{array}
      \right)}\\
  \vec{y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          r^{-1} &
          r^{-1}\,v_\alpha &
          r^{-1}\,v_\delta &
          v_r
        \end{array}
      \right)}
\end{eqnarray}
then the likelihood is
\begin{equation}
  \vec{\hat y} \sim \mathcal{N}(\vec{y}, \mat{C})
\end{equation} where $\mat{C}$ is the covariance matrix, and $\mathcal{N}$ is
normal (Gaussian) distribution.
We adopt a uniform space density prior for the distance and an isotropic
Gaussian for any velocity component, $v$, with a dispersion $\sigma_v=25~\kms$
\begin{eqnarray}
p(r) &=&
  \begin{cases}
    \frac{3}{r_{\rm lim}^3} \, r^2 & \text{if } 0 < r < r_{\rm lim}\\
    0              & \text{otherwise}
  \end{cases}\\
p(v) &=& \frac{1}{\sqrt{2\pi}\,\sigma_v} \,
  \exp\left[-\frac{1}{2} \, \frac{v^2}{\sigma_v^2} \right] \quad .
\end{eqnarray}
%
For each of the two stars, we use \project{emcee}\cite{2013PASP..125..306F}
to generate posterior samples in $(r, v_\alpha, v_\delta, v_r)$
by running 64 walkers for 4608 steps and discarding the first
512 steps as the burn-in period.
For each sample, we convert the heliocentric phase-space coordinates into
Galactocentric coordinates assuming that the Sun's position and velocity are
$\vec x_\odot = (-8.3,\,0,\,0)~{\rm kpc}$ and $\vec v_\odot =
(-11.1,\,244,\,7.25)~\kms$\cite{Schonrich:2010, Schonrich:2012}.

>>>>

% Titles:
% - 90 characters (including spaces)
% - do not normally include numbers, acronyms, abbreviations or punctuation
% - They should include sufficient detail for indexing purposes but be general
%   enough for readers outside the field to appreciate what the paper is about.
\title{
\vspace{-2cm}
  Kronos \& Krios: Evidence for accretion of rocky planets in a comoving pair of G dwarfs
  \vspace{-0.5cm}
}

\author{
  Semyeong Oh$^{1,*}$,\,
  Adrian M. Price-Whelan$^{1}$,\,
  John M. Brewer$^{2}$,\,\\
  David W. Hogg$^{3,4,5,6}$,\,
  David N. Spergel$^{1,3}$,\,
  Justin Myles$^{2}$
}
\date{\small\today}

@@

% affiliations
\vspace{-1cm}
{\small\noindent
  $^1$Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA \\
  $^2$Department of Astronomy, Yale University, 260 Whitney Ave, New Haven, CT 06511, USA \\
  $^3$Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA \\
  $^4$Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA \\
  $^5$Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA \\
  $^6$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg \\
  $^*$To whom correspondence should be addressed: \texttt{semyeong@astro.princeton.edu}
}

@@

% coevality based on stellar parameters
% TODO reference supplement discussion on age here?
Their stellar parameters also indicate that they are coeval with an age of $\sim 4$~Gyr.
We use the distances (inferred from \gaia\ parallaxes), $V$-band magnitudes,
and $B-V$ colors to obtain bolometric luminosities of the two stars\cite{2003AJ....126..778V}.
We then combine the luminosities with $T_\mathrm{eff}$, \elemH{Fe}, \elemH{Si}
in order interpolate the age, mass, and radius of each star
in a grid of Yale-Yonsei model isochrones\cite{2013ApJ...776...87S}.
The best-fit isochrone ages of \sunanalog\ and \bizarreone\ are
$4.00_{-1.56}^{+1.51}$~Gyr and $4.28_{-1.03}^{+1.11}$~Gyr, respectively,
consistent with them being coeval (see Figure~\ref{fig:yyisofitting}).

@@

% high Li
Both stars have high surface \elem{Li} abundances considering
their ages of $\sim 4$~Gyr.
The surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning. %TODO: REFERENCE?
The \elem{Li} doublet at $6707.6$~\AA, which is usually too weak to be
detectable at this age even for high signal-to-noise spectra, is clearly
visible in the spectra of both stars (see \figname~\ref{fig:spec2}).
The measured absolute abundances for \bizarreone\ and \sunanalog\ are
$2.75$~dex and $2.25$~dex, respectively (Myles \etal, in prep),
resulting in the largest abundance difference ($\approx 0.5$~dex) seen
among all measured elements.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{deltaxh_tcrank.pdf}
  \caption{
    Comparing the observed abundance difference vs. \Tcondens\ rank
    to the expected change in solar surface abundance after adding $20$~\mearth\ of
    material with bulk Earth composition (\cite{mcdonough2001composition}).
    The assumed mass fraction in the convective zone is $0.03$.
    All metals are ranked by their \Tcondens\ for solar composition gas,
    and the condensation temperature may be read from the gray line and right y-axis,
    same as in \figname~\ref{fig:relabun_tcrank}.
    Locations of the elements measured for \bizarreone-\sunanalog\ pair are
    indicated by a vertical line and its symbol.
    The close match with the observed abundance difference in \bizarreone-\sunanalog\ pair
    suggests that the abundance difference may be due to accretion of
    $20$~\mearth\ of rocky planetary material.
    The element \elem{Li} is off the plot and indicated with a red arrow
    (see text for details).
  }
  \label{fig:deltaxh_tcrank}
\end{figure}

The abundance differences show a strong increasing trend with condensation
temperature of elements.
In Figure~\ref{fig:deltaxh_tcrank},
we show the abundance difference between \bizarreone\ and \sunanalog\ ordered
by each element's equilibrium \Tcondens\ in solar system composition
gas\citet{2003ApJ...591.1220L}.
The five under-enhanced elements in \bizarreone\
relative to \sunanalog\ are the five most volatile in all elements measured.
The difference in \elem{Mn} ($\Tcondens = 1158$~K) and
\elem{Cr} ($\Tcondens = 1296$~K) suggests a break in $\Tcondens \approx 1200$~K.

We interpret the \Tcondens-dependent metal enhancement of \bizarreone\ relative
to \sunanalog\ and the very high surface \elem{Li} abundance ($A(\elem{Li}) =
2.75$) of \bizarreone\ as evidence for accretion of rocky (refractory-rich)
planetary material in \bizarreone.
In a multi-planet system, dynamical instabilities triggered by planet-planet
scattering\cite{} or encounters with a field star\cite{Malmberg:2011aa} can
lead to planet ejection or accretion.
If one of the two stars in a wide binary accreted more refractory-rich (rocky) planets
than the other, this may imprint a strong \Tcondens\-dependent
abundance differences like those observed in \bizarreone-\sunanalog.

@@

We estimate that accretion of $\approx 15$~\mearth\ of rocky planets
in \bizarreone\ relative to \sunanalog\ can explain the enhancement in
refractory elements as well as the surface Li abundance.
Because the anchor star, \sunanalog,
is very close in its properties (\teff, \logg, \feh, age) to the Sun,
we can use assume the Sun's photosphere abundances\cite{Asplund:2009aa} as the
pre-accretion state of \bizarreone.
In Figure~\ref{fig:deltaxh_tcrank}, we show the expected change in solar abundances,
$\Delta\elemH{X}$, after adding $15$~\mearth\ of bulk Earth
composition\cite{2003TrGeo...2..547M}.
A volatility trend that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known\cite{mcdonough2001composition}.
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding $15$~\mearth\ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggesting that the refractory-enhanced star, \bizarreone\,
accreted $15$~\mearth more of rocky planetary material than \sunanalog.

@@

%TODO: augment this paragraph
% - more geochemical study references
%We stress that while the calculation carried out is useful in
%an order-of-magnitude sense, further investigation of each of the simplifying
%assumptions made is warranted.
%In addition, the composition of bulk Earth has some uncertainties.
%For example, the reported bulk Earth concentration of the siderphile element
%\elem{Mn}, varies from $800$ to $1700$~ppm (\cite{1998psc..book.....L}).
%While the latter value from \cite{mcdonough2001composition} has been used in
%our calculation, the former value would bring the observed $\Delta\elemH{Mn}$
%to an even closer agreement.
%Given these limitations, the level of agreement for $\Delta\elemH{X}$ {\it and}
%\elem{Li} for \bizarreone\ is remarkable.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{yyisofitting.pdf}
  \caption{Probability density function of stellar age, mass, and radius
    from isochrone fitting using Yonsei-Yale isochrones\cite{2013ApJ...776...87S}
    and the star's luminosity, \teff, \feh, and \elemH{Si}.
  }
  \label{fig:yyisofitting}
\end{figure}

@@

Combining the precise radial velocities obtained from spectra with the \tgas\
astrometry, we can compare differences between the inferred 6D phase-space
coordinates of the two stars.
We start by generating posterior samples over the Heliocentric distance, $r$,
tangential velocities, $(v_\alpha, v_\delta)$, and radial velocity, $v_r$,
given the observed parallax, $\hat\pi$, and proper motions,
$(\hat\mu_{\alpha^*}, \hat\mu_\delta)$, and radial velocity, $\hat v_r$.
We assume the noise is Gaussian, and the radial velocity measurement is
uncorrelated with the astrometric measurements.
If we define
\begin{eqnarray}
  \vec{\hat y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          \hat\pi &
          \hat\mu_{\alpha^*} &
          \hat\mu_\delta &
          \hat v_r
        \end{array}
      \right)}\\
  \vec{y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          r^{-1} &
          r^{-1}\,v_\alpha &
          r^{-1}\,v_\delta &
          v_r
        \end{array}
      \right)}
\end{eqnarray}
then the likelihood is
\begin{equation}
  \vec{\hat y} \sim \mathcal{N}(\vec{y}, \mat{C})
\end{equation} where $\mat{C}$ is the covariance matrix, and $\mathcal{N}$ is
normal (Gaussian) distribution.
We adopt a uniform space density prior for the distance and an isotropic
Gaussian for any velocity component, $v$, with a dispersion $\sigma_v=25~\kms$:
\begin{eqnarray}
p(r) &=&
  \begin{cases}
    \frac{3}{r_{\rm lim}^3} \, r^2 & \text{if } 0 < r < r_{\rm lim}\\
    0              & \text{otherwise}
  \end{cases}\\
p(v) &=& \frac{1}{\sqrt{2\pi}\,\sigma_v} \,
  \exp\left[-\frac{1}{2} \, \frac{v^2}{\sigma_v^2} \right] \quad .
\end{eqnarray}
%
For each of the two stars, we use \project{emcee}\cite{2013PASP..125..306F}
to generate posterior samples in $(r, v_\alpha, v_\delta, v_r)$
by running 64 walkers for 4608 steps and discarding the first
512 steps as the burn-in period.
For each sample, we convert the heliocentric phase-space coordinates into
Galactocentric coordinates assuming that the Sun's position and velocity are
$\vec x_\odot = (-8.3,\,0,\,0)~{\rm kpc}$ and $\vec v_\odot =
(-11.1,\,244,\,7.25)~\kms$\cite{Schonrich:2010, Schonrich:2012}.
